---
title: "template_reticulate_ml_rf.Rmd"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Examples for using sklearn in RStudio
### First to test Python and R environment connections
* Load package
```{r}
library(reticulate)
```
* Setup environment for python
```{r}
use_condaenv("deeplearning")
py_config()
```
* Try some data set for connections
```{r}
data(iris)
head(iris)
# write.csv(iris, "iris.csv")
```
* Read csv file from pandas
* Unfortunately, the data 'flower' not shown in R environment through py
```{python}
import pandas as pd
flower = pd.read_csv("iris.csv")
print(flower.head())
```

* Alternatively use 'source_python'
* flower shown in R environment somehow
* data format different
```{r}
source_python('flower.py')
head(py$flower)
```

* Try plotting some data
```{r message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(py$flower, aes(Sepal.Length, Sepal.Width, color=Species)) + 
  geom_point()
```

### Connect 'sklearn' with R
* Build a RandomForestClassifier directly
```{r}
x <- iris
sklearn <- import(module = "sklearn")
rf <- sklearn$ensemble$RandomForestClassifier()
mod <- rf$fit(np_array(as.matrix(x[,1:4])), np_array(as.vector(x[,5])))
mod$feature_importances_
```
* Predict
```{r}
mod$predict_proba(np_array(as.matrix(x[1:2, 1:4])))
```

* Running python script in R
```{r}
matx <- as.matrix(iris)
source_python('rf_mod2.py')
mod2 = classify(matx)
mod2$feature_importances_
```
* model 2 prediction
```{r}
mod2$predict_proba(np_array(as.matrix(x[1:2, 1:4])))
```

* Ref : https://www.youtube.com/watch?v=EJxQSa9lwfM&app=desktop
